\malg: a novel interest management algorithm for distributed simulations of MMOGs

### 2008-07-16 trunk/textos/dsrt/engaoi.tex

Recently, some alternatives to the central server approach have been researched. One of them is to distribute among the own participants the game simulation and the task to broadcast the state updates when they perform actions. Communication between them is in a peer-to-peer manner, forming a decentralized network \cite{schiele2007rpp}. This approach would be the ideal if it did not have not some critical inherent drawbacks. For example, as the player computers participate on the simulation, they need to agree on the resulting state of the game match. If this agreement does not happen, some inconsistency on the game state among the players may occur.

There is another question, concerning the number of transmissions that each peer needs to execute. In the client-server model, each client only needs to send his actions to the server, which processes them and broadcasts the resulting game state to every player. In the peer-to-peer model, each peer becomes responsible for processing its own player's actions and send the resulting state to everyone else to whom that state change might be relevant. The problem is that we cannot assume that every peer has enough bandwidth to do this. Besides, without a central server, the game becomes dependent on the peers' simulation, which can be corrupted in order to result in an invalid state which benefits a player, or even invalidate the whole game session.

@@

This work is based on a virtual environment partitioned into regions, each managed by a different server. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Figure \ref{partitions}). Situations in which two players interacting with each other are connected to different servers imply a higher bandwidth usage, as it is required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of the simulation identical in both servers. Also, the hop count of each message between the clients of these players is increased, since there is one more intermediary.

@@

Before sending the state of an entity to a client, the last transmission time is checked. The next transmission is then scheduled to occur after a certain interval. If the relevance of that state is 1, the normal update interval will be used. If it is less than 1, the normal interval is divided by the relevance. For example, let the normal update interval of a game be \mbox{200 ms}. If the avatar $A_i$, which just sent a state update packet to $A_j$, is at a distance from $A_i$ such that its relevance is 0.5, the next transmission will only happen after an interval of 200 divided by 0.5. So, the player controlling $A_j$ will only receive an update of $A_i$ after 400 ms. Despite this interval is still less than a half second, it represents a reduction of the state update frequency of $A_i$ in 50\%. As they are at a greater distance from one another, and the interval was increased only by 200 ms, this variation will probably be imperceptible by the player who controls $A_j$.

@@

Some works, such as \cite{yu2007nas}, \cite{kim2005tcm} and \cite{svoboda2007taa}, analyse the network traffic generated by large scale games. Based on these analysis, and adopting a conservative posture, the following parameters have been decided to be used in the simulations:

\begin{itemize}
 \item Normal update interval: 250 ms;
 \item State update for one entity: 100 bytes UDP packet;
 \item Duration of each simulated game session: 20 min;
 \item Virtual environment area: 750 x 750;
 \item View distance: 120;
 \item Critical distance: 40;
 \item View angle: 180°.
\end{itemize}

>>>>

Recently, some alternatives to the central server approach have been researched. One of them is to distribute among the own participants the game simulation and the task to broadcast the state updates when they perform actions. The communication between them is peer-to-peer, forming a decentralized network \cite{schiele2007rpp}. This approach would be the ideal if it did not have some critical inherent drawbacks. For example, as the player computers participate on the simulation, they need to agree on the resulting state of the game match. If this agreement does not happen, some inconsistency on the game state among the players may occur.

There is another question, concerning the number of transmissions that each peer needs to execute. In the client-server model, each player only needs to send his actions to the server, which processes them and broadcasts the resulting game state to the other players. In the peer-to-peer model, each peer becomes responsible for processing its own player's actions and send the resulting state to everyone else to whom that state change is relevant. The problem is that we cannot assume that every peer has enough bandwidth to do this. Besides, without a central server, the game becomes dependent on the peers' simulation, which may be corrupted in order to result in an invalid state that benefits a player, or even invalidate the whole game session.

@@

This work is part of the P2PSE \cite{crippa2007pps} project, which is currently based on a peer-to-peer distribution, but still looking for an ideal model. Hence, we propose here an alternative that consists of a virtual environment partitioned into regions - each managed by a different server - on which the \malg\ algorithm is based. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Figure \ref{partitions}).

One issue that relates to that type of partitioning of the virtual environment concerns the borders between regions. If an avatar is close to the border between two regions, it will be necessary some exchange of information between the servers which manage them. This information consists of state updates of the entities that are interacting with each other despite being located in different regions. Such situations imply a higher bandwidth usage, as it is also required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of the simulation identical among all of them. Furthermore, the hop count of each message between the clients will be increased, since there will be two intermediary servers.

@@

Before sending the state of an entity to a client, the last transmission time is checked. The next transmission is then scheduled to occur after a certain interval. If the relevance of that state is 1, the normal update interval will be used. If it is less than 1, the normal interval is divided by the relevance. For example, let the normal update interval of a game be \mbox{200 ms}. Also, let $A_i$ be the avatar of a player who just received a state update packet from $A_j$. If $A_j$ is at a distance from $A_i$ such that its relevance is 0.5, the next transmission will only happen after an interval of 200 divided by 0.5. Therefore, the player controlling $A_i$ will only receive another update from $A_j$ after 400 ms. Despite this interval is still less than a half second, it represents a reduction of the state update frequency of $A_j$ in 50\%, to the player of $A_i$. As they are at a greater distance from one another, and the interval was increased only by 200 ms, this variation will probably be imperceptible by the player.

@@

Some works, such as \cite{yu2007nas}, \cite{kim2005tcm} and \cite{svoboda2007taa}, analyse the network traffic generated by large scale games. Based on these analyses, and adopting a conservative posture, the following parameters have been decided to be used in the simulations:

\begin{itemize}
 \item Normal update interval: 250 ms;
 \item State update for one entity: 100 bytes UDP packet;
 \item Duration of each simulated game session: 20 min;
 \item Virtual environment area: 750 $\times$ 750 area units;
 \item View distance: 120 length units;
 \item Critical distance: 40 length units;
 \item View angle: 180°.
\end{itemize}
